Growing Up with Nell: A Narrative Interface for Literacy

### 2012-03-12 paper.tex
Unabbreviate "Figure" -- we've got the space.
Turn off authors for submission, so that IDC can do their blind reviewing thing.
Copy-editing the text at 4:30am.  Hope I'm getting everything.

% THIS IS SIGPROC-SP.TEX - VERSION 3.1
% WORKS WITH V3.2SP OF ACM_PROC_ARTICLE-SP.CLS
%
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission,
% you need to 'insert'  your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%\documentclass{acm_proc_article-sp}
\documentclass[preprint]{sig-alternate}
\usepackage{url}
\usepackage{ifthen}\newboolean{noauthors}\setboolean{noauthors}{false}

@@

We also seek to eliminate the ``ceiling'' between the children and
the authors of the Nell system itself.  Nell should be capable of
teaching about itself, and a capable child can dive into Nell's source
code and make meaningful changes.  In section~\ref{sec:turtles} we
describe how this goal is approached.

@@

The constructionist learning philosophy emphasizes creation and
tangible interaction.  Nell uses direct interaction on a tablet
computer to maintain the child's connection to their work
(Fig.~\ref{fig:xo3}).

The direct interaction model lowers the floor by eliminating the need
to learn an abstract touchpad or mouse interface.  It also supports
our literacy goals by allowing direct handwriting instruction and the
use of handwriting as an interface.

\section{Technology}
Our implementation choices further our goals of modularity and
decentralization.  Nell's implementation is open source and uses
standard web technologies to broaden the contributor base.  In this
section we describe the technologies that enable Nell's four key ideas.

@@

Story modules can be inherited and extended, as shown in the figure.
This makes it easy to take an existing story and modify it to
better suit particular interests, a particular type of learner---or
just to add variety.  This can lead to conflicts: multiple versions of
a story, or scenes within a story, may have their
preconditions satisfied simultaneously.

Conflicts are resolved by consulting a \textit{group} associated with
each scene.  Each group has a default priority and names a scene to be
invoked when a conflict exists; the actions associated with that scene
will (eventually) alter the priorities of the conflicting scenes to
resolve the conflict.  The resolution scene may include any number of
actions.  It may select randomly, or based a selection on
child preferences, curriculum progress, or inferred learning style.
The resolution scene may even initiate a new multi-scene story.

For example, the \textit{story-intro-group}
used in Figure~\ref{fig:rules} defaults to a low priority so that
continuing a story in progress is preferred over starting a new story.
When the story in progress is complete, several new story openers will be in
conflict.  The resolution scene for \textit{story-intro-group} brings us
to the controls of a magic airship which flies between locations
corresponding to the various story continuations.  Eventually the
child decides to fly to Mathland to hear the next episode in her
agent Mathis' adventure, and that episode's priority is bumped to
allow us to transition to it without further conflict.

We plan to write an integrated story editor for Nell.  Our model will
be the Wide Ruled authoring tool, which was successfully used by
non-technical story creators~\cite{skorupski:2009}.

@@

Our implementation extends AIML~\cite{aiml:2005},
the markup language developed for the three-time Loebner Prize--winning
AliceBot.%
\footnote{Wilcox identifies several shortcomings with the expressive
  power of AIML, proposing ChatScript as a
  replacement~\cite{wilcox:2010}.  AIML's community and
  diversity of third-party implementations make it our choice at present.}
Each story module can include AIML fragments extending the
conversational capabilities of Nell's agents.

Figure~\ref{fig:aiml}
shows an AIML fragment augmenting an agent with
commands suitable for a text adventure story module.
In such a story the child might write instructions such as
\texttt{go north}, \texttt{go east}, or \texttt{open door} in
addition to usual agent dialog.
These new commands fire events which are referenced by scene
preconditions.  For example, \texttt{north} may
trigger a scene which causes the activity to draw a new location on
the page, causes the agent to write and speak a description of the new
location, and which shifts the AIML topic to enable additional vocabulary
suitable for the new location.

@@

We have developed a block-oriented view of JavaScript, which we call
TurtleScript~\cite{turtlescript}.  This allows us to use
the TurtleScript view of JavaScript as the single
implementation language for Nell.  TurtleScript scales from Logo-like
pedagogical tasks (in subsets of the language) up to the
implementation of TurtleScript and the
Nell system itself.  TurtleScript's block view is isomorphic to
JavaScript's standard textual
representation to preserve the code-reading skills of existing
programmers (Figure~\ref{fig:turtlescript}).  We have begun implementation of a
tablet programming environment for TurtleScript which uses
drag-and-drop, direct interaction, and intelligent prompting to reduce
text entry requirements.

@@

TurtleScript allows the child to dig as far into Nell's implementation
as their capabilities allow.  Although the operating system and the
browser implementation remain inaccessible, incorporating a
meta-circular interpreter for JavaScript (such as \cite{narcissus}) allows us
to expose the system all the way down to the programming
language implementation,
% Via TurtleScript, this code is all visible in a
% friendly block-based form.

@@

The TinkRBook system~\cite{chang:tinkrbook} shares a similar focus on
personalization and literacy.  Nell attempts to scale these
ideas with a modular extensible narrative system and a broader range
of pedagogic material.

The Sugar Learning Platform~\cite{sugar} is an earlier effort at
child education.  We've learned directly from this effort.  Nell
focuses on children even further from educational infrastructure.
The narrative at Nell's core attempts to provide students additional
pedagogical guidance, and the direct interaction model avoids keyboard
problems which have plagued Sugar's initial deployments.  Nell also
attempts to provide a greater sense of ownership through pervasive
customization.

@@

Automatic narrative generation has seen attention in
gaming, simulation, and training contexts~\cite{riedl:2008}.  Nell's
use of these ideas in a pedagogical context is unique.

\section{Conclusions}
We have described the design of a novel narrative direct-interface
system for teaching literacy, which is personalized for and grows with
its child owner.  Personalized direct interface engages the child, and
narrative guides them through the pedagogic material.  The low-cost
solar-powered hardware allows Nell to reach further into the
least-developed areas of the world, to help those far from traditional
educational infrastructure.

>>>>

% THIS IS SIGPROC-SP.TEX - VERSION 3.1
% WORKS WITH V3.2SP OF ACM_PROC_ARTICLE-SP.CLS
%
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission,
% you need to 'insert'  your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%\documentclass{acm_proc_article-sp}
\documentclass[preprint]{sig-alternate}
\usepackage{url}
\usepackage{ifthen}\newboolean{noauthors}\setboolean{noauthors}{true}

@@

We also seek to eliminate the ``ceiling'' between the children and
the authors of the Nell system itself.  Nell should be capable of
teaching about itself, and a capable child can dive into Nell's source
code and make meaningful changes.
%In Section~\ref{sec:turtles} we describe how this goal is approached.

@@

The constructionist learning philosophy emphasizes creation and
tangible interaction.  Nell uses direct interaction on a tablet
computer to maintain the child's connection to their work
(Figure~\ref{fig:xo3}).  In particular, Nell uses handwriting
recognition as a primary interface.

The direct interaction model lowers the floor by eliminating the need
to learn an abstract touchpad or mouse interface.  It also supports
our literacy goals by allowing direct handwriting instruction and
development of motor skills.

\section{Technology}
Our technology choices further our goals of modularity and
decentralization.  Nell's implementation is free and open source and uses
standard web technologies to broaden the contributor base.  In this
section we describe the technologies that enable Nell's four key ideas.

@@

Story modules can be inherited and extended.
This makes it easy to take an existing story and modify it to
better suit particular interests, a particular type of learner---or
just to add variety.  This can lead to conflicts: multiple versions of
a story, or scenes within a story, may have their
preconditions satisfied simultaneously.

Conflicts are resolved by consulting a \textit{group} associated with
each scene.  Each group has a default priority and names a scene to be
invoked when a conflict exists; the actions associated with that scene
will (eventually) alter the priorities of the conflicting scenes to
resolve the conflict.  The resolution scene may include any number of
actions.  It may select randomly among the conflicting scenes or base
its selection on
child preferences, curriculum progress, or inferred learning style.
The resolution scene may even initiate a new multi-scene story.

For example, the \textit{story-intro-group}
used in Figure~\ref{fig:rules} defaults to a low priority so that
continuing a story in progress is preferred over starting a new story.
When a story module is completed, several new story openings will be in
conflict.  The resolution scene for \textit{story-intro-group} brings us
to the controls of a magic airship which flies between locations
corresponding to the various story continuations.  In that set of
scenes, the child may eventually decide to fly to Mathland to hear the
next episode in her agent Mathis' adventure, which would bump up the
appropriate episode's priority to allow us to make the transition
without further conflict.

Nell will also include an integrated story editor.  Our exemplar is
the Wide Ruled authoring tool, which was successfully used by
non-technical story creators~\cite{skorupski:2009}.

@@

Our implementation extends AIML~\cite{aiml:2005},
the markup language developed for the Loebner Prize--winning
AliceBot.%
\footnote{Wilcox identifies several shortcomings with the expressive
  power of AIML, proposing ChatScript as a
  replacement~\cite{wilcox:2010}.  AIML's community and
  diversity of third-party implementations make it our choice at present.}
Each story module can include AIML fragments extending the
conversational capabilities of Nell's agents.

Figure~\ref{fig:aiml}
shows an AIML fragment augmenting an agent with
commands suitable for a text adventure story module.
In such a story the child might write instructions such as
\texttt{go north}, \texttt{go east}, or \texttt{open door} in
addition to the usual dialog with the agent.
These new commands fire events which are referenced by scene
preconditions.  For example, \texttt{north} may
trigger a scene which causes the activity to draw a new location on
the page, causes the agent to write and speak a description of the new
location, and which causes a shift in the AIML \textit{topic} to
enable additional vocabulary suitable for the new location.

@@

We have developed a block-oriented view of JavaScript, which we call
TurtleScript~\cite{turtlescript}.
The TurtleScript view of JavaScript is the single
implementation language for Nell.  TurtleScript scales from Logo-like
pedagogical tasks (in subsets of the language) up to the
implementation of TurtleScript and the
Nell system itself.  As Figure~\ref{fig:turtlescript} shows,
TurtleScript's block view is isomorphic to
JavaScript's standard textual
representation to preserve the code-reading skills of existing
programmers.  We have begun implementation of a
tablet programming environment for TurtleScript which uses
drag-and-drop, direct interaction, and intelligent prompting to reduce
text entry requirements.

@@

TurtleScript allows the child to dig as far into Nell's implementation
as their capabilities allow.  Although the operating system and the
browser implementation remain inaccessible, incorporating a
meta-circular interpreter for JavaScript (such as \cite{narcissus}) allows us
to expose the system all the way down to the programming
language implementation.
% Via TurtleScript, this code is all visible in a
% friendly block-based form.

@@

The TinkRBook system~\cite{chang:tinkrbook} shares a similar focus on
personalization and literacy.  Nell attempts to scale up these
ideas with a modular extensible narrative system and a broader range
of pedagogic material.

The Sugar Learning Platform~\cite{sugar} is an earlier effort at
child education which we've been directly involved in.  Nell's focus is
on children even further from educational infrastructure than those
targeted by Sugar.
The narrative at Nell's core attempts to provide students additional
pedagogical guidance, and the direct interaction model avoids keyboard
problems which have plagued Sugar's initial deployments.  Nell also
attempts to provide a greater sense of ownership through pervasive
customization.

@@

Automatic narrative generation has seen attention in
gaming, simulation, and training contexts~\cite{riedl:2008}.
We believe
Nell's use of these ideas in a pedagogical context is unique.

\section{Conclusions}
We have described the design of a novel narrative direct-interface
system for teaching literacy, which is personalized for and grows with
its child owner.  Personalized direct interface engages the child, and
narrative guides them through the pedagogic material.  The low-cost
solar-powered deployment hardware allows Nell to reach further into the
least-developed areas of the world to help those without traditional
educational infrastructure.
